{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-03-10 thesis/education.tex
fix incomplete references/labels/citations, move some chapters to their own files

%for a more compact document, add the option openany to avoid
%starting all chapters on odd numbered pages
\documentclass[12pt]{cmuthesis}

% This is a template for a CMU thesis.  It is 18 pages without any content :-)
% The source for this is pulled from a variety of sources and people.
% Here's a partial list of people who may or may have not contributed:
%
%        bnoble   = Brian Noble
%        caruana  = Rich Caruana
%        colohan  = Chris Colohan
%        comar    = Cyrus Omar
%        jab      = Justin Boyan
%        josullvn = Joseph O'Sullivan
%        jrs      = Jonathan Shewchuk
%        kosak    = Corey Kosak
%        mjz      = Matt Zekauskas (mattz@cs)
%        pdinda   = Peter Dinda
%        pfr      = Patrick Riley
%        dkoes = David Koes (me)

% My main contribution is putting everything into a single class files and small
% template since I prefer this to some complicated sprawling directory tree with
% makefiles.

% some useful packages
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[numbers,sort]{natbib}
\usepackage[backref,pageanchor=true,plainpages=false, pdfpagelabels, bookmarks,bookmarksnumbered,
%pdfborder=0 0 0,  %removes outlines around hyper links in online display
]{hyperref}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}

%% custom fonts

\usepackage{fontspec}
\usepackage{xeCJK}

\setmainfont{XCharter}
%\setmainfont{Utopia}[
%	Path = ./fonts/,
%	UprightFont = UtopiaStd-Regular.otf,
%	BoldFont = UtopiaStd-Semibold.otf,
%	ItalicFont = UtopiaStd-Italic.otf,
%	BoldItalicFont = UtopiaStd-SemiboldIt.otf
%	]

\setCJKmainfont{HeiseiMinStd-W5}[Path = ./fonts/]

% Approximately 1" margins, more space on binding side
%\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar]{geometry}
%for general printing (not binding)
\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar,hmarginratio=1:1]{geometry}

% Provides a draft mark at the top of the document.
\draftstamp{\today}{DRAFT}

\begin{document}
\frontmatter

%initialize page style, so contents come out right (see bot) -mjz
\pagestyle{empty}

\title{ %% {\it \huge Thesis Proposal}\\
{\bf Practical Concurrency Testing} \\
\normalsize \vspace{1em}
\begin{tabular}{c}
{\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}
\end{tabular}}
\author{Ben Blum}
\date{Decembruary 2018}
\Year{2018}
\trnumber{}

\committee{
Garth Gibson, Chair \\
David A. Eckhardt \\
Brandon Lucia \\
Haryadi Gunawi, University of Chicago
}

\support{}
\disclaimer{}

% copyright notice generated automatically from Year and author.
% permission added if \permission{} given.

\keywords{landslide terminal, baggage claim, ground transportation, ticketing}

\maketitle

\begin{dedication}
For Elsie, Margie, Johnny, Mel, Eve, and David.
%, who brought me here today.
\end{dedication}

\pagestyle{plain} % for toc, was empty

%% Obviously, it's probably a good idea to break the various sections of your thesis
%% into different files and input them into this file...

\begin{abstract}
Concurrent programming presents a challenge to students and experts alike because of the complexity of multithreaded interactions and the difficulty to reproduce and reason about bugs.
Stateless model checking is a concurrency testing approach which forces a program to interleave its threads in many different ways, checking for bugs each time.
This technique is powerful, in principle capable of finding any nondeterministic bug in finite time,
but suffers from exponential explosion as program size increases.
Checking an exponential number of thread interleavings is not a practical or predictable approach for programmers to find concurrency bugs before their project deadlines.

In this thesis, I develop several new techniques to make stateless model checking more practical for human use.
I have built Landslide, a stateless model checker specializing in undergraduate operating systems class projects.
Landslide extends the traditional model checking algorithm with a new framework for automatically managing
%the exploration of
multiple state spaces according to their estimated completion times,
which I show quickly finds bugs should they exist and also quickly verifies correctness otherwise.
I evaluate Landslide's suitability for inexpert use by presenting the results of many semesters providing it to students in 15-410, CMU's Operating System Design and Implementation class,
and more recently, students in similar classes at the University of Chicago and Berkeley.
%Finally, I will present several new techniques that allow stateless model checking to be practically employed on real-world programs.
%Finally, I will explore broader impact by extending Landslide to test some real-world programs and to be used by students at other universities.
Finally, I extend Landslide with a new concurrency model for hardware transactional memory,
and evaluate several real-world transactional benchmarks to show that stateless model checking can keep up with the developing concurrency demands of real-world programs.
\end{abstract}

\begin{acknowledgments}
% things to thank
% stimhack
% csd bgames
% undergrad frenz
% new frenz
% 410 ta community
% emarhavil
%
% \\

I thank my dog, Louie.
\end{acknowledgments}

\tableofcontents
% TODO: figure out how to make latex abbreviate these
\listoffigures
\listoftables

\mainmatter

%% Double space document for easy review:
%\renewcommand{\baselinestretch}{1.66}\normalsize

% The other requirements Catherine has:
%
%  - avoid large margins.  She wants the thesis to use fewer pages,
%    especially if it requires colour printing.
%
%  - The thesis should be formatted for double-sided printing.  This
%    means that all chapters, acknowledgements, table of contents, etc.
%    should start on odd numbered (right facing) pages.
%
%  - You need to use the department standard tech report title page.  I
%    have tried to ensure that the title page here conforms to this
%    standard.
%
%  - Use a nice serif font, such as Times Roman.  Sans serif looks bad.
%
% Other than that, just make it look good...

\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,192}

\newtheorem{theorem}{Theorem}

\newcommand\llama[1]{\includegraphics[width=#1]{llama.pdf}}
\newcommand\llitem{\item[\llama{1.2em}]}

\newcommand{\inspirationallinebreak}{\vspace{0.25em}}
\newcommand{\inspirationalhyphen}{-\hspace{-0.15em}-\xspace}
\newcommand\inspirationalquote[2]{\begin{flushright}
	\vspace{-1em}
	{
	\fontspec[Path=./fonts/]{AlexaStd}
	{\em #1}
	\inspirationallinebreak

	-\hspace{-0.15em}-\hspace{-0.15em}-{#2}
	}
	\vspace{2em}
\end{flushright}}

\input{intro}
\input{overview}
\input{landslide}

\chapter{Quicksand}
\inspirationalquote{
% original
% {\footnotesize 嫌な事も悲しい事もあったけど、守りたい物だって沢山この世界にはあったから…} \\
% trans. slightly paraphrased from crunchyroll subtitle; removed "..."s, added past tense for あった
There were awful, sad things in this world.
But there were a lot of things worth protecting, too.
}
{Kaname Madoka, Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}
%\inspirationalquote{Always, somewhere, someone is fighting for you.
%As long as you remember her, you are not alone.}
%{Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}

\chapter{Education}
\inspirationalquote{Knowing the students might one day
%find a way to
fix their concurrency bugs...
it fills you with determination.}{Undertale (paraphrased)}

\chapter{Transactions}
%\inspirationalquote{

\chapter{Related Work}
\inspirationalquote{
\begin{tabular}{p{0.58\textwidth}}
To test if your paper makes a genuine contribution to its discipline,
see if you can afford a generous tone in the "Related Work" section.
\end{tabular}}
{Conor McBride}

\chapter{Future Work}
% TODO: ask for permission for this one
% https://twitter.com/Poem4your_sprog/status/878747092788838400
\inspirationalquote{
\begin{tabular}{p{0.54\textwidth}}
A little bit of being kind,
a tiny open door; \\
A nicer slice of peace of mind
begets a little more. \\
It won't correct the world tonight,
nor change tomorrow too; \\
But maybe if you do it right,
you'll find it changes *you*.
\end{tabular}}
{Sam "Poem\_for\_your\_sprog" Garland}

\chapter{Conclusion}
\inspirationalquote{
	\begin{tabular}{p{0.84\textwidth}}
	"If you are a god, Zeus, as the stories claim, then why did you create evolution?
	Why did you make a world that can only grow through cruelty and pain?"
	\\
	For a moment\inspirationalhyphen{}surely this meant death was near\inspirationalhyphen{}she thought she heard him answer:
	"My child, I was shaped by the gods that came before me, as you were shaped by me.
	The choice I had was between creation and oblivion, life and death.
	And I chose life, because any life is better than no life,
	because as long as there is life, there is hope\inspirationalhyphen{}if not for us, then for some generation to come."
	\\
	"Then how are you a god, if you can offer so little?" she whispered, feeling death creep closer.
	\\
	"I am a god because I take upon myself the burden of creation," the statue replied.
	\\
	"Then we are all gods," Alexandra said, and pushed the button.
	\end{tabular}}
%{Galatea, The Talos Principle}
{The Talos Principle}

asdfasdasdf
%\appendix
%\include{appendix}

\backmatter

%\renewcommand{\baselinestretch}{1.0}\normalsize

% By default \bibsection is \chapter*, but we really want this to show
% up in the table of contents and pdf bookmarks.
\renewcommand{\bibsection}{\chapter{\bibname}
\inspirationalquote{Only a doctor of philosophy, Darth.}{Robert Marsh}
}
%\renewcommand{\bibpreamble}{This text goes between the ``Bibliography''
%  header and the actual list of references}
\bibliographystyle{plainnat} % plain n'at
\bibliography{citations}

\end{document}

>>>>

\chapter{Education}
\label{chap:education}

\inspirationalquote{Knowing the students might one day
%find a way to
fix their concurrency bugs...
it fills you with determination.}{Undertale (paraphrased)}

% TODO

### 2018-03-10 thesis/intro.tex
fix incomplete references/labels/citations, move some chapters to their own files

\begin{enumerate}
	\item {\bf Meaningful state spaces (Chapter \ref{chap:quicksand}).}
		I will present {\em Iterative Deepening}, a new algorithm for navigating the trade-off in how many preemption points to test at once.
		Iterative Deepening incorporates state space estimation \cite{estimation} to decide on-the-fly whether each state space is worth pursuing, and uses data race analysis \cite{tsan} to find new preemption point candidates based on a program's dynamic behaviour.
		This section will include a large evaluation of the technique, comparing its performance to three prior work approaches across 600+ unique tests.
		I will show that Iterative Deepening of preemption points outperforms prior work in terms both of finding bugs quickly and of completely verifying correctness when no bug exists.
	\item {\bf Educational use (Chapters \ref{chap:410} and \ref{chap:pintos}).}
		For the past five semesters, I have offered a fully-automated version of Landslide to students in 15-410, CMU's undergraduate Operating System Design and Implementation class \cite{kspec,thrlib}, for use as a debugging aid during the thread library project.
		Recently I have also extended Landslide to handle Pintos kernel projects from other universities \cite{pintos}.
		In the two most recent semesters, I collaborated with Operating Systems course staff at two such schools, the University of Chicago and University of California at Berkeley,
		to provide debugging feedback to their students.

>>>>

\begin{enumerate}
	\item {\bf Meaningful state spaces (Chapter \ref{chap:quicksand}).}
		I will present {\em Iterative Deepening}, a new algorithm for navigating the trade-off in how many preemption points to test at once.
		Iterative Deepening incorporates state space estimation \cite{estimation} to decide on-the-fly whether each state space is worth pursuing, and uses data race analysis \cite{tsan} to find new preemption point candidates based on a program's dynamic behaviour.
		This section will include a large evaluation of the technique, comparing its performance to three prior work approaches across 600+ unique tests.
		I will show that Iterative Deepening of preemption points outperforms prior work in terms both of finding bugs quickly and of completely verifying correctness when no bug exists.
	\item {\bf Educational use (Chapter \ref{chap:education}).}
		For the past five semesters, I have offered a fully-automated version of Landslide to students in 15-410, CMU's undergraduate Operating System Design and Implementation class \cite{kspec,thrlib}, for use as a debugging aid during the thread library project.
		Recently I have also extended Landslide to handle Pintos kernel projects from other universities \cite{pintos}.
		In the two most recent semesters, I collaborated with Operating Systems course staff at two such schools, the University of Chicago and University of California at Berkeley,
		to provide debugging feedback to their students.

### 2018-03-10 thesis/landslide.tex
fix incomplete references/labels/citations, move some chapters to their own files

\chapter{Landslide}
\inspirationalquote{Somewhere is the promise of an uncharted trail, with 700 branching limbs and 700 ways to fail.}
{ThouShaltNot, "Cardinal Directions"}

@@

This section describes the features of Landslide the average student user should expect to interact with.
% TODO: make more specific section reference
Separate user guides also exist, described in Chapter~\ref{chap:410}.

@@

\begin{itemize}
	\item Basic statistics such as the current instruction pointer, thread ID,
		stack trace of current thread at the moment of preemption,
		depth in the tree, parent node pointer, etc.;
	\item Snapshots of the current state of the scheduler (\sect{\ref{sec:landslide-scheduler}})
		and memory accesses and heaps (\sect{\ref{sec:landslide-memory}});
	\item Simulator-dependent data needed to time travel and resume execution
		from this checkpoint (\sect{\ref{sec:landslide-timetravel}});
	\item List of parent/ancestor nodes with memory conflicts and/or happens-before edges to this one
		for DPOR (\sect{\ref{sec:landslide-dpor}});
	\item Current estimated state space proportion and execution time for the subtree rooted at this node
		(not necessarily fully explored yet) for estimation \sect{\ref{sec:landslide-estimate}};
	\item Whether this point is an {\tt xbegin} invocation
		and if so what {\tt xabort} codes are possible and/or already explored for this transaction
		(\sect{\ref{chap:txn}}).
\end{itemize}

@@

{\tt memory.c} is responsible for all manner of memory access analysis.
It tracks heap allocations, checks reads and writes in the heap region against same;
tracks reads and writes (in any region) from each thread
and checks them against each other
for DPOR (\sect{\ref{sec:landslide-dpor}})
and data race analysis (\sect{\ref{sect:landslide-datarace}}).

@@

\subsection{Pintos-specific features}
\label{sec:landlside-pintosspecifics}

>>>>

\chapter{Landslide}
\label{chap:landslide}

\inspirationalquote{Somewhere is the promise of an uncharted trail, with 700 branching limbs and 700 ways to fail.}
{ThouShaltNot, "Cardinal Directions"}

@@

This section describes the features of Landslide the average student user should expect to interact with.
% TODO: make more specific section reference
Separate user guides also exist, described in Chapter~\ref{chap:education}.

@@

\begin{itemize}
	\item Basic statistics such as the current instruction pointer, thread ID,
		stack trace of current thread at the moment of preemption,
		depth in the tree, parent node pointer, etc.;
	\item Snapshots of the current state of the scheduler (\sect{\ref{sec:landslide-scheduler}})
		and memory accesses and heaps (\sect{\ref{sec:landslide-memory}});
	\item Simulator-dependent data needed to time travel and resume execution
		from this checkpoint (\sect{\ref{sec:landslide-timetravel}});
	\item List of parent/ancestor nodes with memory conflicts and/or happens-before edges to this one
		for DPOR (\sect{\ref{sec:landslide-dpor}});
	\item Current estimated state space proportion and execution time for the subtree rooted at this node
		(not necessarily fully explored yet) for estimation \sect{\ref{sec:landslide-estimate}};
	\item Whether this point is an {\tt xbegin} invocation
		and if so what {\tt xabort} codes are possible and/or already explored for this transaction
		(\sect{\ref{chap:tm}}).
\end{itemize}

@@

{\tt memory.c} is responsible for all manner of memory access analysis.
It tracks heap allocations, checks reads and writes in the heap region against same;
tracks reads and writes (in any region) from each thread
and checks them against each other
for DPOR (\sect{\ref{sec:landslide-dpor}})
and data race analysis (\sect{\ref{sec:landslide-datarace}}).

@@

\subsection{Pintos-specific features}
\label{sec:landslide-pintosspecifics}

### 2018-03-10 thesis/quicksand.tex
fix incomplete references/labels/citations, move some chapters to their own files

%for a more compact document, add the option openany to avoid
%starting all chapters on odd numbered pages
\documentclass[12pt]{cmuthesis}

% This is a template for a CMU thesis.  It is 18 pages without any content :-)
% The source for this is pulled from a variety of sources and people.
% Here's a partial list of people who may or may have not contributed:
%
%        bnoble   = Brian Noble
%        caruana  = Rich Caruana
%        colohan  = Chris Colohan
%        comar    = Cyrus Omar
%        jab      = Justin Boyan
%        josullvn = Joseph O'Sullivan
%        jrs      = Jonathan Shewchuk
%        kosak    = Corey Kosak
%        mjz      = Matt Zekauskas (mattz@cs)
%        pdinda   = Peter Dinda
%        pfr      = Patrick Riley
%        dkoes = David Koes (me)

% My main contribution is putting everything into a single class files and small
% template since I prefer this to some complicated sprawling directory tree with
% makefiles.

% some useful packages
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[numbers,sort]{natbib}
\usepackage[backref,pageanchor=true,plainpages=false, pdfpagelabels, bookmarks,bookmarksnumbered,
%pdfborder=0 0 0,  %removes outlines around hyper links in online display
]{hyperref}
\usepackage{subfigure}
\usepackage{xspace}
\usepackage[linesnumbered,ruled]{algorithm2e}
\usepackage{algpseudocode}

%% custom fonts

\usepackage{fontspec}
\usepackage{xeCJK}

\setmainfont{XCharter}
%\setmainfont{Utopia}[
%	Path = ./fonts/,
%	UprightFont = UtopiaStd-Regular.otf,
%	BoldFont = UtopiaStd-Semibold.otf,
%	ItalicFont = UtopiaStd-Italic.otf,
%	BoldItalicFont = UtopiaStd-SemiboldIt.otf
%	]

\setCJKmainfont{HeiseiMinStd-W5}[Path = ./fonts/]

% Approximately 1" margins, more space on binding side
%\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar]{geometry}
%for general printing (not binding)
\usepackage[letterpaper,twoside,vscale=.8,hscale=.75,nomarginpar,hmarginratio=1:1]{geometry}

% Provides a draft mark at the top of the document.
\draftstamp{\today}{DRAFT}

\begin{document}
\frontmatter

%initialize page style, so contents come out right (see bot) -mjz
\pagestyle{empty}

\title{ %% {\it \huge Thesis Proposal}\\
{\bf Practical Concurrency Testing} \\
\normalsize \vspace{1em}
\begin{tabular}{c}
{\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}
\end{tabular}}
\author{Ben Blum}
\date{Decembruary 2018}
\Year{2018}
\trnumber{}

\committee{
Garth Gibson, Chair \\
David A. Eckhardt \\
Brandon Lucia \\
Haryadi Gunawi, University of Chicago
}

\support{}
\disclaimer{}

% copyright notice generated automatically from Year and author.
% permission added if \permission{} given.

\keywords{landslide terminal, baggage claim, ground transportation, ticketing}

\maketitle

\begin{dedication}
For Elsie, Margie, Johnny, Mel, Eve, and David.
%, who brought me here today.
\end{dedication}

\pagestyle{plain} % for toc, was empty

%% Obviously, it's probably a good idea to break the various sections of your thesis
%% into different files and input them into this file...

\begin{abstract}
Concurrent programming presents a challenge to students and experts alike because of the complexity of multithreaded interactions and the difficulty to reproduce and reason about bugs.
Stateless model checking is a concurrency testing approach which forces a program to interleave its threads in many different ways, checking for bugs each time.
This technique is powerful, in principle capable of finding any nondeterministic bug in finite time,
but suffers from exponential explosion as program size increases.
Checking an exponential number of thread interleavings is not a practical or predictable approach for programmers to find concurrency bugs before their project deadlines.

In this thesis, I develop several new techniques to make stateless model checking more practical for human use.
I have built Landslide, a stateless model checker specializing in undergraduate operating systems class projects.
Landslide extends the traditional model checking algorithm with a new framework for automatically managing
%the exploration of
multiple state spaces according to their estimated completion times,
which I show quickly finds bugs should they exist and also quickly verifies correctness otherwise.
I evaluate Landslide's suitability for inexpert use by presenting the results of many semesters providing it to students in 15-410, CMU's Operating System Design and Implementation class,
and more recently, students in similar classes at the University of Chicago and Berkeley.
%Finally, I will present several new techniques that allow stateless model checking to be practically employed on real-world programs.
%Finally, I will explore broader impact by extending Landslide to test some real-world programs and to be used by students at other universities.
Finally, I extend Landslide with a new concurrency model for hardware transactional memory,
and evaluate several real-world transactional benchmarks to show that stateless model checking can keep up with the developing concurrency demands of real-world programs.
\end{abstract}

\begin{acknowledgments}
% things to thank
% stimhack
% csd bgames
% undergrad frenz
% new frenz
% 410 ta community
% emarhavil
%
% \\

I thank my dog, Louie.
\end{acknowledgments}

\tableofcontents
% TODO: figure out how to make latex abbreviate these
\listoffigures
\listoftables

\mainmatter

%% Double space document for easy review:
%\renewcommand{\baselinestretch}{1.66}\normalsize

% The other requirements Catherine has:
%
%  - avoid large margins.  She wants the thesis to use fewer pages,
%    especially if it requires colour printing.
%
%  - The thesis should be formatted for double-sided printing.  This
%    means that all chapters, acknowledgements, table of contents, etc.
%    should start on odd numbered (right facing) pages.
%
%  - You need to use the department standard tech report title page.  I
%    have tried to ensure that the title page here conforms to this
%    standard.
%
%  - Use a nice serif font, such as Times Roman.  Sans serif looks bad.
%
% Other than that, just make it look good...

\newcommand\simics{\textsc{Simics}}
\newcommand{\sect}[1]{\S #1}
\newcommand\hilight[2]{\color{#1}#2\color{black}}
\definecolor{orange}{RGB}{192,96,0}
\definecolor{olivegreen}{RGB}{0,127,0}
\definecolor{brickred}{RGB}{192,0,0}
\definecolor{commentblue}{RGB}{0,0,192}

\newtheorem{theorem}{Theorem}

\newcommand\llama[1]{\includegraphics[width=#1]{llama.pdf}}
\newcommand\llitem{\item[\llama{1.2em}]}

\newcommand{\inspirationallinebreak}{\vspace{0.25em}}
\newcommand{\inspirationalhyphen}{-\hspace{-0.15em}-\xspace}
\newcommand\inspirationalquote[2]{\begin{flushright}
	\vspace{-1em}
	{
	\fontspec[Path=./fonts/]{AlexaStd}
	{\em #1}
	\inspirationallinebreak

	-\hspace{-0.15em}-\hspace{-0.15em}-{#2}
	}
	\vspace{2em}
\end{flushright}}

\input{intro}
\input{overview}
\input{landslide}

\chapter{Quicksand}
\inspirationalquote{
% original
% {\footnotesize 嫌な事も悲しい事もあったけど、守りたい物だって沢山この世界にはあったから…} \\
% trans. slightly paraphrased from crunchyroll subtitle; removed "..."s, added past tense for あった
There were awful, sad things in this world.
But there were a lot of things worth protecting, too.
}
{Kaname Madoka, Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}
%\inspirationalquote{Always, somewhere, someone is fighting for you.
%As long as you remember her, you are not alone.}
%{Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}

\chapter{Education}
\inspirationalquote{Knowing the students might one day
%find a way to
fix their concurrency bugs...
it fills you with determination.}{Undertale (paraphrased)}

\chapter{Transactions}
%\inspirationalquote{

\chapter{Related Work}
\inspirationalquote{
\begin{tabular}{p{0.58\textwidth}}
To test if your paper makes a genuine contribution to its discipline,
see if you can afford a generous tone in the "Related Work" section.
\end{tabular}}
{Conor McBride}

\chapter{Future Work}
% TODO: ask for permission for this one
% https://twitter.com/Poem4your_sprog/status/878747092788838400
\inspirationalquote{
\begin{tabular}{p{0.54\textwidth}}
A little bit of being kind,
a tiny open door; \\
A nicer slice of peace of mind
begets a little more. \\
It won't correct the world tonight,
nor change tomorrow too; \\
But maybe if you do it right,
you'll find it changes *you*.
\end{tabular}}
{Sam "Poem\_for\_your\_sprog" Garland}

\chapter{Conclusion}
\inspirationalquote{
	\begin{tabular}{p{0.84\textwidth}}
	"If you are a god, Zeus, as the stories claim, then why did you create evolution?
	Why did you make a world that can only grow through cruelty and pain?"
	\\
	For a moment\inspirationalhyphen{}surely this meant death was near\inspirationalhyphen{}she thought she heard him answer:
	"My child, I was shaped by the gods that came before me, as you were shaped by me.
	The choice I had was between creation and oblivion, life and death.
	And I chose life, because any life is better than no life,
	because as long as there is life, there is hope\inspirationalhyphen{}if not for us, then for some generation to come."
	\\
	"Then how are you a god, if you can offer so little?" she whispered, feeling death creep closer.
	\\
	"I am a god because I take upon myself the burden of creation," the statue replied.
	\\
	"Then we are all gods," Alexandra said, and pushed the button.
	\end{tabular}}
%{Galatea, The Talos Principle}
{The Talos Principle}

asdfasdasdf
%\appendix
%\include{appendix}

\backmatter

%\renewcommand{\baselinestretch}{1.0}\normalsize

% By default \bibsection is \chapter*, but we really want this to show
% up in the table of contents and pdf bookmarks.
\renewcommand{\bibsection}{\chapter{\bibname}
\inspirationalquote{Only a doctor of philosophy, Darth.}{Robert Marsh}
}
%\renewcommand{\bibpreamble}{This text goes between the ``Bibliography''
%  header and the actual list of references}
\bibliographystyle{plainnat} % plain n'at
\bibliography{citations}

\end{document}

>>>>

\chapter{Quicksand}
\label{chap:quicksand}
\inspirationalquote{
% original
% {\footnotesize 嫌な事も悲しい事もあったけど、守りたい物だって沢山この世界にはあったから…} \\
% trans. slightly paraphrased from crunchyroll subtitle; removed "..."s, added past tense for あった
There were awful, sad things in this world.
But there were a lot of things worth protecting, too.
}
{Kaname Madoka, Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}
%\inspirationalquote{Always, somewhere, someone is fighting for you.
%As long as you remember her, you are not alone.}
%{Mahou Shoujo Madoka{\raisebox{0.1em}{$\scriptstyle \bigstar$}}Magica}

% TODO

\section{Preemption points} % ??? -- label needed from reference to LAST_CALL in landslide.tex
\label{sec:quicksand-pps}

\section{Evaluation}
\label{sec:quicksand-eval}
